Learning constraints in tabular data

### 2016-05-13 Paper/ECAI/tabular.tex
added preliminary figure

\newcommand{\eccalc}[2]{\ensuremath{#1 = #2}}
\newcommand{\ecrank}[2]{\eccalc{#1}{\mathit{RANK}(#2)}}
\newcommand{\ecfkey}[2]{\ensuremath{\mathit{FOREIGNKEY}(#1,#2)}}
\newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
\newcommand{\eclookupf}[4]{\ensuremath{\mathit{LOOKUP}_{\mathit{#4}}(#1, #2, #3)}}
\newcommand{\eclookup}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{}}}
\newcommand{\eclookupprod}[5]{\eccalc{#1}{#2 \times \eclookupf{#3}{#4}{#5}{}}}
\newcommand{\eclookupfuzzy}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{fuzzy}}}
\newcommand{\ecperm}[1]{\ensuremath{\mathit{PERMUTATION}(#1)}}
\newcommand{\ecseries}[1]{\ensuremath{\mathit{SERIES}(#1)}}
\newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
\newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
\newcommand{\ectotal}[3]{\eccalc{#1}{\mathit{PREV}(#1) + #2 - #3}}
\newcommand{\ecproj}[2]{\eccalc{#1}{\mathit{PROJECT}(#2)}}
\newcommand{\ecaggc}[3]{\eccalc{#2}{\mathit{#1}(#3, col)}}
\newcommand{\ecaggr}[3]{\eccalc{#2}{\mathit{#1}(#3, row)}}
\newcommand{\ecsumc}[2]{\eccalc{#1}{\mathit{SUM}(#2, col)}}
\newcommand{\ecsumr}[2]{\eccalc{#1}{\mathit{SUM}(#2, row)}}
\newcommand{\ecaggif}[5]{\eccalc{#2}{\mathit{#1IF}(#3, #4, #5)}}
\newcommand{\ecsumif}[4]{\eccalc{#1}{\mathit{SUMIF}(#2, #3, #4)}}
\newcommand{\ecsumprod}[3]{\eccalc{#1}{\mathit{SUMPRODUCT}(#2, #3)}}

@@

  \begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo.png}
  \end{center}
  \vspace{-10pt}
  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations.}
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.30\textwidth}
  {\small
    \begin{align*}
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
%
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggc{AVERAGE}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{MAX}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{MIN}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggif{SUM}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{MAX}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints that are satisfied in the case study (Figure on the left) and expected to be found by our implementation. Functional constraint are omitted.}
  \label{fig:sol_example}
\end{subfigure}
\end{figure*}

@@

\subsection{Autocompletion}
Autocompletion can be seen as using knowledge derived from data at time~$t_0$ and new input data at time~$t_1 > t_0$ to predict data that the user has not yet written.
Therefore, constraints are learned at time~$t_0$ and combined with data from a new row~$i$ (column~$j$) that the user has added to a table~$T$ by time~$t_1$.
Constraints that do not conflict with the added data are considered \textit{viable}.
Viable constraints that have enough input data to compute the result for a blank cell~$T[i,j]$ in the new row (column) are \textit{active}, which means they can be used to predict the outcome for that cell.
\textit{Active} constraints can be used to autocomplete blank cells such that the user does not have to fill in the values manually.
Once the user has stopped editing the new row~$i$ (column~$j$), the table~$T$ is updated to include~$i$ ($j$) and constraints are learned once again.

\begin{figure*}[thb]
  \begin{center}
    \includegraphics[width=0.85\textwidth]{figures/fbi_figure_highlighted.png}
  \end{center}
  \caption{Real world tabular constraint reconstruction: FBI crime statistics}
  \label{fig:fbi}
\end{figure*}

\subsection{Error detection}
We consider two cases of error detection.
The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
However, instead of considering viable constraints, we look at \textit{conflict} constraints, i.e. constraints that were learned at time~$t_0$ but do not agree with the newly added data at time~$t_1$.
Such conflict constraints can indicate errors in the input.
Since conflict constraints could also be accidental constraints that are correctly invalidated by the new data, the size of the data at time ~$t_0$ should be large enough.

>>>>

\newcommand{\eccalc}[2]{\ensuremath{#1 = #2}}
\newcommand{\ecrank}[2]{\eccalc{#1}{\textit{RANK}(#2)}}
\newcommand{\ecfkey}[2]{\ensuremath{\textit{FOREIGNKEY}(#1,#2)}}
\newcommand{\ecalldiff}[1]{\ensuremath{\textit{ALLDIFFERENT}(#1)}}
\newcommand{\eclookupf}[4]{\ensuremath{\textit{LOOKUP}_{\textit{#4}}(#1, #2, #3)}}
\newcommand{\eclookup}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{}}}
\newcommand{\eclookupprod}[5]{\eccalc{#1}{#2 \times \eclookupf{#3}{#4}{#5}{}}}
\newcommand{\eclookupfuzzy}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{fuzzy}}}
\newcommand{\ecperm}[1]{\ensuremath{\textit{PERMUTATION}(#1)}}
\newcommand{\ecseries}[1]{\ensuremath{\textit{SERIES}(#1)}}
\newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
\newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
\newcommand{\ectotal}[3]{\eccalc{#1}{\textit{PREV}(#1) + #2 - #3}}
\newcommand{\ecproj}[2]{\eccalc{#1}{\textit{PROJECT}(#2)}}
\newcommand{\ecaggc}[3]{\eccalc{#2}{\textit{#1}(#3, col)}}
\newcommand{\ecaggr}[3]{\eccalc{#2}{\textit{#1}(#3, row)}}
\newcommand{\ecsumc}[2]{\eccalc{#1}{\textit{SUM}(#2, col)}}
\newcommand{\ecsumr}[2]{\eccalc{#1}{\textit{SUM}(#2, row)}}
\newcommand{\ecaggif}[5]{\eccalc{#2}{\textit{#1IF}(#3, #4, #5)}}
\newcommand{\ecsumif}[4]{\eccalc{#1}{\textit{SUMIF}(#2, #3, #4)}}
\newcommand{\ecsumprod}[3]{\eccalc{#1}{\textit{SUMPRODUCT}(#2, #3)}}

@@

  \begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo.png}
  \end{center}
  \vspace{-10pt}
  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations.}
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.30\textwidth}
  {\small
    \begin{align*}
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
%
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints that are satisfied in the case study (Figure on the left) and expected to be found by our implementation. Functional constraint are omitted.}
  \label{fig:sol_example}
\end{subfigure}
\end{figure*}

@@

\subsection{Autocompletion}
\begin{figure}[thb]
  \begin{center}
    \includegraphics[width=0.33\textwidth]{figures/autocompletion_example.png}
  \end{center}
  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
  \label{fig:autocompletion_example}
\end{figure}

Autocompletion can be seen as using knowledge derived from data at time~$t_0$ and new input data at time~$t_1 > t_0$ to predict data that the user has not yet written.
Therefore, constraints are learned at time~$t_0$ and combined with data from a new row~$i$ (column~$j$) that the user has added to a table~$T$ by time~$t_1$.
Constraints that do not conflict with the added data are considered \textit{viable}.
Viable constraints that have enough input data to compute the result for a blank cell~$T[i,j]$ in the new row (column) are \textit{active}, which means they can be used to predict the outcome for that cell.
\textit{Active} constraints can be used to autocomplete blank cells such that the user does not have to fill in the values manually.
Once the user has stopped editing the new row~$i$ (column~$j$), the table~$T$ is updated to include~$i$ ($j$) and constraints are learned once again.

Let us illustrate this on one of the spreadsheets from the gathered dataset. In Figure \ref{fig:autocompletion_example}, \sergey{to be finished}

\subsection{Error detection}
\begin{figure*}[thb]
  \begin{center}
    \includegraphics[width=0.85\textwidth]{figures/fbi_figure_highlighted.png}
  \end{center}
  \caption{Real world tabular constraint reconstruction: FBI crime statistics}
  \label{fig:fbi}
\end{figure*}

We consider two cases of error detection.
The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
However, instead of considering viable constraints, we look at \textit{conflict} constraints, i.e. constraints that were learned at time~$t_0$ but do not agree with the newly added data at time~$t_1$.
Such conflict constraints can indicate errors in the input.
Since conflict constraints could also be accidental constraints that are correctly invalidated by the new data, the size of the data at time ~$t_0$ should be large enough.
